---
title: "Part_I_explore_exercise"
output: html_notebook
---

```{r}
# Load Libraries

library(tidyverse)


```

**3.2.4 Exercises**

**1.Run ggplot(data = mpg). What do you see?**

```{r}
ggplot(data = mpg)
```
**Ans:** I see a blank canvas.

**2. How many rows are in mpg? How many columns?**

```{r}
dim(mpg)
```
**Ans:** Rows: 234, Coloumn: 11

**3. What does the drv variable describe? Read the help for ?mpg to find out.**

**Ans:** f = front-wheel drive, r = rear wheel drive, 4 = 4wd

**4. Make a scatterplot of hwy vs cyl.**

**Ans:** 
```{r}
ggplot(data = mpg, aes(x = cyl, y = hwy)) + geom_point()

```

**5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?**

**Ans:** 

```{r}
ggplot(data = mpg, aes(x = class, y = drv)) + geom_point()
```

Both variables are factors. 

**3.3.1 Exercises**

1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Ans: The code should be,

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
str(mpg)
```

Ans: Categorical:manufacturer, model, trans, drv ,fl, class ; Continuous: cty, hwy, displ

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

Ans: 
```{r}
ggplot(data = mpg, aes(x= hwy, y = cty, color = fl, size = cty, shape = drv)) + geom_point()
```

4. What happens if you map the same variable to multiple aesthetics?

Ans: 
```{r}
ggplot(data = mpg, aes(x= hwy, y = cyl, color = drv, size = drv, shape = drv)) + geom_point()
```

5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

Ans:For shapes that have a border (like 21), you can colour the inside and outside separately. Use the stroke aesthetic to modify the width of the border

```{r}
ggplot(data = mpg, aes(x= hwy, y = cty)) + geom_point(color = "red", size = 10, shape = 6, stroke = 2)

```

6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(data = mpg, aes(x= displ, y = hwy, colour = class >= "subcompact")) + geom_point()

```


# 3.5.1 Exercises


1. What happens if you facet on a continuous variable?

There is one facet for each value e.g. a facet on displ.

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

The facets are empty when there is no data for the according combination e.g. rear wheel drive (r) with 4 or 5 cylinder is not listed. The 7 cylinder factes are missed entirely.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(~ hwy)

```

3. What plots does the following code make? What does . do?

The “attribute ~ dot” notation plots the attribute values without a column attribute, thus showing multiple row-wise plots for each attribute value. The y-axis is repeated. With “dot ~ attribute” the row attribute is missing, thus showing column-wise the plots. Then the x-axis is repeated.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


4. Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

With faceting it is easier to examine the indivual classes. With coloring it is easier to see how the classes are clustered overall. With larger datasets it’s more likely that you want to see the overall clustering instead of the individual point clouds.


```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + facet_wrap(~ class, nrow = 2)
```


```{r}
ggplot(data = mpg, aes(x = displ, y = hwy, color = class)) + geom_point()
```

5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

```{r}
?facet_wrap
```

facet_wrap wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid() because most displays are roughly rectangular.

Facet grids do not have these configuration because the rows and cols are determined by the specified attributes.


nrow, ncol:    Number of rows and columns.

scales:    should Scales be fixed ("fixed", the default), free ("free"), or free in one dimension ("free_x", "free_y").

shrink:    If TRUE, will shrink scales to fit output of statistics, not raw data. If FALSE, will be range of raw data before statistical summary.


6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

When putting the more levels on the row axis, then the y-axis would shrink so that it is harder to see which actual values are at the points as shown in the plot.

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + facet_grid(class ~ drv)
```


































